Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-08-04 ltx/data_paper/DataPaper.tex
minor edit to introduction

The human brain brain is arguably the most complicated system known to man. The development of a complete theory for its function is one of the greatest challenges faced by scientists today. To address this challenge, researchers have naturally used both theoretical and experimental frameworks to develop and test hypotheses investigating brain function. Experimental frameworks are typically designed to uncover causal relationships between the systems properties or parameters and its function (reverse engineering the brain). Theoretical studies are typically used in one of two ways. The first is to explain data acquired in an experiment and the second is to predict system behaviour.

@@

By assimilating experimental or clinical measurements with mathematical models one can infer unmeasured or latent system properties or parameters. Inference of patient-specific parameters has the the potential to revolutionise the treatment of disease. Typically, one can measure the blood-oxygen level dependent signal through fMRI or the electromagnetic fields of the brain through EEG or MEG. In clinical neurology, these measurement modalities give a kind or a `where and when' indication of the presence of a pathology, but provide little information on the causative mechanisms for disease. A classic example of this is in epilepsy management, where the disease is diagnosed using EEG. If electrical seizures are recorded, a medication plan is prescribed. The choice of medication is not based on the data acquired from the EEG, but based on the clinician's experience with similar cases and sometimes trial and error because similar electrographic seizure can have fundamentally different mechanisms of initiation (ref JTs paper). Ideally in epilepsy monitoring we would like to image the concentration of ions, the synaptic dynamics, the connectivity strength and structure etc. to better inform a treatment plan by understand physiological changes that lead to seizures. This is also highly desirable for epileptic seizure prediction..

>>>>

The human brain is arguably the most complicated system known to man. The development of a complete theory for its function is one of the greatest challenges faced by scientists today. To address this challenge, researchers have naturally used both theoretical and experimental frameworks to develop and test hypotheses investigating brain function. Experimental frameworks are typically designed to uncover causal relationships between the systems properties or parameters and its function (reverse engineering the brain). Theoretical studies are typically used in one of two ways. The first is to explain data acquired in an experiment and the second is to predict system behaviour.

@@

By assimilating experimental or clinical measurements with mathematical models one can infer unmeasured or latent system properties or parameters. Inference of patient-specific parameters has the potential to revolutionise the treatment of disease. Typically, one can measure the blood-oxygen level dependent signal through fMRI or the electromagnetic fields of the brain through EEG or MEG. In clinical neurology, these measurement modalities give a kind or a `where and when' indication of the presence of a pathology, but provide little information on the causative mechanisms for disease. A classic example of this is in epilepsy management, where the disease is diagnosed using EEG. If electrical seizures are recorded, a medication plan is prescribed. The choice of medication is not based on the data acquired from the EEG, but based on the clinician's experience with similar cases and sometimes trial and error because similar electrographic seizure can have fundamentally different mechanisms of initiation (ref JTs paper). Ideally in epilepsy monitoring we would like to image the concentration of ions, the synaptic dynamics, the connectivity strength and structure etc. to better inform a treatment plan by understand physiological changes that lead to seizures. This is also highly desirable for epileptic seizure prediction..
